\bf Excuse me, do you have a moment to talk about version control?

### 2017-10-05 response-to-review.Rmd

---
title: "Response to review"
output: html_document
---

I am currently unable to get back into the Scholar One TAS system, since my link has expired. TAS editor Hadley Wickham was able to forward me a marked up PDF and I have based my edits on the annotations and comments I see there.

I made all the small copy edits that I could make sense of. I was 100% sure what some of the suggestions were, however, so a few loci remain 'as is'. I think it is possible that Preview doesn't display all the mark up as intended? I did my best.

The series of changes that I made can be seen as a sequence of commits in this branch on the repository in which I've written this article:

https://github.com/dsscollection/git-github-for-stats/commits/TAS

Below are a few specific notes about changes (or non-changes).

Added citations for the iris data.

> The flow of the paper is somewhat disrupted by repeated use of bullet points.  Is there a way to fold some of these into complete paragraphs?  (This may not be one of the ones to modify, but do search out instances throughout where you can collapse these lists to have better text flow.)

I converted every set of bullet points to a paragraph that I could. Some I have left because I feel being set off or presented as individual steps is actually conducive to understanding.

> Replace "deltas" with "differences" or "incremental changes" (etc.)

"Deltas" are actually a technical term in Git. There are such things as "delta-chains", which describe all the incremental changes necessary to transform one version of a repo to another. I use this word intentionally. I think it's best to keep it. Finally, "delta" is used in alot in math to describe differences, so I suspect this audience will understand it.

> This last part may be extraneous detail that our readers don't need to know.

This comment is about SHA-1 checksums. Again, one doesn't have to do much Googling about Git to be confronted with this vocabulary and these concepts. I feel like the goal of this article is to demystify Git and it's best to introduce the key terms.

### 2017-10-05 README.tex
Successfully creating PDFs for TAS

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{listings}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}

@@

@@

Git has been re-purposed by the data science community
\citep{Ram2013, git-for-humans, ten-simple-rules-git}. We use it to
manage the motley collection of files that make up typical data
analytical projects, which consist of data, figures, reports, and, yes,
source code. Even those who identify more as statistician than data
scientist generally have a similar mix of files that are the artifacts
of a project.

A lone ranger, working on a single computer, can benefit from adopting
version control. But not nearly enough to justify the pain of
installation and workflow upheaval. There are much easier ways to get
versioned back ups of files, if that's all you're worried about.

@@

Many people who don't use Git unwittingly re-invent a poor man's version
of it. Figure \ref{fig:diy-vs-git} depicts a hypothetical analysis of
the iris data, captured in a single R source file. With informal version
control, contributors create derivative copies of \texttt{iris.R},
decorating the file name with initials, dates, and other descriptors.
Even when working alone, this leads to multiple versions of
\texttt{iris.R} of indeterminate relatedness (Figure
\ref{fig:diy-vs-git}A). In collaborative settings based on email
distribution, the original file swiftly becomes part of a complicated
phylogeny that no amount of ``Track changes'' and good intentions can
resolve Figure \ref{fig:diy-vs-git}B).

@@

The Git way is to track the evolution of \texttt{iris.R}, through a
series of commits, each equipped with an explanatory message. Figure
\ref{fig:diy-vs-git}C depicts this linear, \emph{in situ} process, with
development and time flowing from bottom to top. Figure
\ref{fig:diy-vs-git}D shows the same history for a common collaborative
Git workflow, where contributors work independently but sync regularly
to a common version. Especially important versions get a human-readable
tag, e.g., ``draft-01'', to signal a meaningful milestone. Yes, there is
some pain in adopting the formalism of Git, but it is worth it.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  When you are new with Git (or, frankly, even when you're not), it's
  common to goof up the Git infrastructure for a project. Note that your
  files can be intact and safe, even while the Git tracking is a bit
  confused. Of course there are official Git remedies, but sometimes the
  easiest fix is to clone a fresh copy from GitHub, patch things up with
  the changes that only exist locally, and move on with your life. This
  workaround obviously requires the existence of a recent copy on
  GitHub.
\item
  The highly functional web interfaces mentioned above are often the
  most pleasant and natural way to navigate and search your files, even
  though all the same information exists locally. It is a pleasure to
  browse through your own work, across multiple projects or files and
  across time, as if it's a well-designed website. You must push your
  work to GitHub to enjoy this.
\end{enumerate}

@@

Issues can be assigned to specific people and they can be labelled, e.g.
``bug'', ``simulation-study'', or ``final-exam''. Coupled with the
ability to cross-link issues and the project files or file changes, you
have extraordinary power to document why things have happened in the
past and to organize what needs to happen in the future.

@@

If I've convinced you to experiment with Git and GitHub, you need to do
some initial setup. These first steps happen once or, for some steps,
once per computer. This is excerpted from
\href{http://happygitwithr.com}{Happy Git and GitHub for the useR},
which holds battle-tested instructions honed over several years in
\href{http://stat545.com}{STAT 545} at the University of British
Columbia.

@@

Once this setup is done, you are ready to start using Git and GitHub
with your projects. Some general recommendations for agony reduction:

\begin{itemize}
\tightlist
\item
  I repeat: consider using a graphical front-end for Git, a.k.a. a Git
  client, versus restricting yourself to the command line interface.
\item
  Establish confidence in the basics (e.g.~make a change, commit it,
  push it) before wading into more advanced usage (e.g.~branching).
\item
  Commit yourself to Git usage on a project that will provide sustained
  practice over several months. Usage in a course is great, because it
  provides a relentless stream of small deadlines.
\item
  Realize that no one is giving out Git style points. It's OK to
  ``power-cycle'', i.e.~re-initialize the Git repository, to get
  unstuck.
\end{itemize}

@@

Here is the daily workflow:

\begin{itemize}
\tightlist
\item
  Go about your usual business, writing R scripts or authoring reports
  in LaTeX or R Markdown. But instead of only \emph{saving} individual
  files, periodically you make a \textbf{commit}, which takes a snapshot
  of all the files in the entire project.

  \begin{itemize}
  \tightlist
  \item
    If you have ever versioned a file
    \href{http://www.phdcomics.com/comics/archive.php?comicid=1531}{by
    adding your initials or the date}, you have effectively made a
    commit, albeit only for a single file. It is a version that is
    significant to you and that you might want to inspect or revert to
    later.
  \end{itemize}
\item
  Push commits to GitHub periodically.

  \begin{itemize}
  \tightlist
  \item
    This is like sharing a document with colleagues on DropBox or
    sending it out as an email attachment. By pushing to GitHub, you
    make your work and all your accumulated progress accessible to
    others.
  \end{itemize}
\end{itemize}

@@

\section{Markdown is special on
GitHub}\label{markdown-is-special-on-github}

This may seem unrelated to Git, GitHub, and R, but it is now necessary
to talk about
\href{https://daringfireball.net/projects/markdown/syntax}{Markdown}.
Markdown is a markup language, like HTML and LaTeX, but designed to be
as lightweight as possible. The goal is still to separate form and
content, but also to prioritize human-readability, even at the cost of
fancy features. Markdown is in wide use on sites like
\href{https://en.support.wordpress.com/markdown/}{WordPress},
\href{https://stackoverflow.com/editing-help}{StackOverflow}, and, yes,
\href{https://help.github.com/categories/writing-on-github/}{GitHub}.
These sites use Markdown because it allows a diverse population of site
users to author decent-looking web content, with hyperlinks and some
formatting. Do not build this up into some heroic, LaTeX-level learning
task, for it is not. If you can write an email, you can write Markdown.

@@

\section{Markdown is special for R
users}\label{markdown-is-special-for-r-users}

Markdown is especially holy for R users because of
\href{http://rmarkdown.rstudio.com}{R Markdown}, which is just Markdown
that includes chunks of R code. Figure \ref{fig:report-from-r-or-rmd}A
shows a simple \texttt{.Rmd} document for our iris example. Again, do
not regard R Markdown as something you must clear your schedule to
learn. If you can write email and a bit of R code, you can write R
Markdown. The
\href{https://CRAN.R-project.org/package=rmarkdown}{rmarkdown package}
\citep{rmd-pkg} converts R Markdown (\texttt{.Rmd} files) to Markdown
(\texttt{.md} files), running the code and inserting the results,
including figures, into the document. This is powered by another
package, \href{https://CRAN.R-project.org/package=knitr}{knitr}
\citep{knitr-pkg, knitr-book}, under the hood. This process is made
especially easy in RStudio, but is by no means limited to users of that
application. Any R user can call \texttt{rmarkdown::render("foo.Rmd")}.

@@

These R-derived Markdown files, if committed and pushed, then enjoy the
usual privileged treatment on GitHub already described above. Once an
\texttt{.Rmd} file has been rendered to \texttt{.md}, anyone viewing it
on GitHub can read the prose, study the R code, \textbf{and view the
results of running that code}, including figures. Figure
\ref{fig:report-from-r-or-rmd}C shows how the \texttt{.Rmd} document
from Figure \ref{fig:report-from-r-or-rmd}A looks when rendered to
\texttt{.md} and pushed to GitHub. It is the best of all worlds, because
the code is revealed and, by definition, is the code that produced the
results. And yet a reader can gaze upon the product in a web browser,
without needing to download the code, install all necessary
dependencies, and run it.

>>>>

% !TeX program = pdfLaTeX
\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{textcomp}
\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}

@@

%% load any required packages here

@@

Git has been re-purposed by the data science community
\citep{Ram2013, git-for-humans, ten-simple-rules-git}. We use it to
manage the motley collection of files that make up typical data
analytical projects, which consist of data, figures, reports, and source
code. Even those who identify more as statistician than data scientist
generally have a similar mix of files that are the artifacts of a
project.

A lone ranger, working on a single computer, can benefit from adopting
version control, But not nearly enough to justify the pain of
installation and workflow upheaval. There are much easier ways to get
versioned back ups of files, if that's all you're worried about.

@@

Many people who don't use Git unwittingly re-invent a poor man's version
of it. Figure \ref{fig:diy-vs-git} depicts a hypothetical analysis of
the iris data \citep[\citet{anderson}]{fisher}, captured in a single R
source file. With informal version control, contributors create
derivative copies of \texttt{iris.R}, decorating the file name with
initials, dates, and other descriptors. Even when working alone, this
leads to multiple versions of \texttt{iris.R} of indeterminate
relatedness (Figure \ref{fig:diy-vs-git}A). In collaborative settings
based on email distribution, the original file swiftly becomes part of a
complicated phylogeny that no amount of ``Track changes'' and good
intentions can resolve (Figure \ref{fig:diy-vs-git}B).

@@

The Git way is to track the evolution of \texttt{iris.R}, through a
series of commits, each equipped with an explanatory message. Figure
\ref{fig:diy-vs-git}C depicts this linear, \emph{in situ} process, with
development and time flowing from bottom to top. Figure
\ref{fig:diy-vs-git}D shows the same history for a common collaborative
Git workflow, where contributors work independently but sync regularly
to a common version. Especially important versions get a human-readable
tag, e.g., ``draft-01'', to signal a meaningful milestone. There is some
pain in adopting the formalism of Git, but it is worth it.

@@

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  When you are new with Git (or, frankly, even when you're not), it's
  common to damage the Git infrastructure for a project. Note that your
  files can be intact and safe, even while the Git tracking is a bit
  confused. Of course there are official Git remedies, but sometimes the
  easiest fix is to clone a fresh copy from GitHub, patch things up with
  the changes that only exist locally, and move on with your life. This
  workaround obviously requires the existence of a recent copy on
  GitHub.
\item
  The highly functional web interfaces mentioned above are often the
  most pleasant and natural way to navigate and search your files, even
  though all the same information exists locally. It is a pleasure to
  browse through your own work, across multiple projects or files and
  across time, as if it's a well-designed website. You must push your
  work to GitHub to enjoy this.
\end{enumerate}

@@

Issues can be assigned to specific people and they can be labelled, e.g.
``bug'', ``simulation-study'', or ``final-exam''. Coupled with the
ability to cross-link issues and the project files or file changes, you
have the power to document why things have happened in the past and to
organize what needs to happen in the future.

@@

If I've convinced you to experiment with Git and GitHub, you need to do
some initial setup. These first steps happen once or, for some steps,
once per computer. These instructions are excerpted from
\href{http://happygitwithr.com}{Happy Git and GitHub for the useR},
which holds battle-tested instructions honed over several years in
\href{http://stat545.com}{STAT 545} at the University of British
Columbia.

@@

Once this setup is done, you are ready to start using Git and GitHub
with your projects. Here are some general recommendations for agony
reduction. As suggested above, it may be worth using a graphical
front-end for Git, a.k.a. a Git client, versus restricting yourself to
the command line interface. It is fruitful to establish confidence in
basic operations (e.g.~make a change, commit it, push it) before wading
into more advanced usage (e.g.~branching). It is helpful to begin with a
project that will provide sustained practice over several months. For
example, usage in a course is great, because it provides a relentless
stream of small deadlines. Finally, it is liberating to realize that no
one is giving out Git style points. It is acceptable to ``power-cycle'',
i.e.~re-initialize the Git repository, to get unstuck.

@@

The daily workflow is probably not dramatically different from what you
do currently. You work in the usual way, writing R scripts or authoring
reports in LaTeX or R Markdown. But instead of only \emph{saving}
individual files, periodically you make a \textbf{commit}, which takes a
snapshot of all the files in the entire project. If you have ever
versioned a file
\href{http://www.phdcomics.com/comics/archive.php?comicid=1531}{by
adding your initials or the date}, you have effectively made a commit,
albeit only for a single file. It is a version that is significant to
you and that you might want to inspect or revert to later. Periodically,
you push commits to GitHub. This is like sharing a document with
colleagues on DropBox or sending it out as an email attachment. By
pushing to GitHub, you make your work and all your accumulated progress
accessible to others.

@@

\section{Markdown is special on GitHub and for R
users}\label{markdown-is-special-on-github-and-for-r-users}

This may seem unrelated to Git, GitHub, and R, but it is now necessary
to talk about
\href{https://daringfireball.net/projects/markdown/syntax}{Markdown}.
Markdown is a markup language, like HTML and LaTeX, but designed to be
as lightweight as possible. The goal is still to separate form and
content, but also to prioritize human-readability, even at the cost of
fancy features. Markdown is in wide use on sites such as
\href{https://en.support.wordpress.com/markdown/}{WordPress},
\href{https://stackoverflow.com/editing-help}{Stack Overflow}, and
\href{https://help.github.com/categories/writing-on-github/}{GitHub}.
These sites use Markdown because it allows a diverse population of site
users to create decent-looking web content, with hyperlinks and some
formatting. Do not build this up into some heroic, LaTeX-level learning
task, for it is not. If you can write an email, you can write Markdown.

@@

Markdown is special for R users because of
\href{http://rmarkdown.rstudio.com}{R Markdown}, which is just Markdown
that includes chunks of R code. Figure \ref{fig:report-from-r-or-rmd}A
shows a simple \texttt{.Rmd} document for our iris example. Again, do
not regard R Markdown as something you must clear your schedule to
learn. If you can write email and a bit of R code, you can write R
Markdown. The
\href{https://CRAN.R-project.org/package=rmarkdown}{rmarkdown package}
\citep{rmd-pkg} converts R Markdown (\texttt{.Rmd} files) to Markdown
(\texttt{.md} files), running the code and inserting the results,
including figures, into the document. This is powered by another
package, \href{https://CRAN.R-project.org/package=knitr}{knitr}
\citep{knitr-pkg, knitr-book}, under the hood. This process is made
especially easy in RStudio, but is by no means limited to users of that
application. Any R user can call \texttt{rmarkdown::render("foo.Rmd")}.

@@

These R-derived Markdown files, if committed and pushed, then enjoy the
usual privileged treatment on GitHub already described above. Once an
\texttt{.Rmd} file has been rendered to \texttt{.md}, anyone viewing it
on GitHub can read the prose, study the R code, and view the results of
running that code, including figures. Figure
\ref{fig:report-from-r-or-rmd}C shows how the \texttt{.Rmd} document
from Figure \ref{fig:report-from-r-or-rmd}A looks when rendered to
\texttt{.md} and pushed to GitHub. It is the best of all worlds, because
the code is revealed and, by definition, is the code that produced the
results. And yet a reader can gaze upon the product in a web browser,
without needing to download the code, install all necessary
dependencies, and run it.
